---
title: "Exchange Rate Forecasting Using Bayesian VARs Model"
author: "Qingqing Pang"

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This research report explores how Bayesian VARs model predict AUD/USD exchange rate.
> **Keywords.** Bayesian Vars, Exchange rate

# Objective and Motivation

The objective of this research is to use the Bayesian Vector Autoregressions (VARs) method to forecast the exchange rate of the US dollar exchange rate against the Australian dollar.

The ability to accurately forecast the foreign exchange rate is crucial for Australia’s global trade and investment. Given the prominence of the US dollar as the world's primary reserve currency, the monetary policies of the US Federal Reserve have a worldwide effect on the world economy, its significant influence on the Australia currency market should be important for domestic investors and policy maker as it can directly impact the AUD/USD exchange rate. Besides, America as one of the major trading partners for Australia, can impact bilateral trade flows and eventually affect the value of the AUD. Apart from the external from foreign countries, the domestic economic indicators can also be one of the determinants of the AUD/USD exchange rate.

The research is aimed to address the question for example, how the AUD/USD exchange rate will be in 3 months or even longer 1 year?

# Data and Variables

To better forecast the change in the exchange rate, the 13 variables are selected as follows which contain both domestic and US economic indicators that affect the exchange rate in different ways.

Real GDP and interest rates have a significant effect on the exchange rate, a higher realGDP and interest rate in Australia may increase the demand for AUD, which will lead to an appreciation of AUD and a rise in the AUD/USD exchange rate. A higher CPI indicates a lower purchasing power relative to foreign currency which may lead to a depreciation of the domestic currency. The unemployment rate can in some way represent business activity and a country with a high unemployment rate will lower the attractiveness for foreign investors and weaken the domestic currency competitiveness in the currency market. The balance of trade, which is the difference between exports and imports, also can influence the demand for its currency, a trade surplus in AUD may increase the demand for AUD dollar. 

  -   $erate_{t}$: AUD/USD exchange rate

* AUS economic indicators
  +   $crate\_au_{t}$: The Cash Rate Target, Australia
  +   $rgdp\_au_{t}$: The Real Gross Domestic Product, Australia
  +   $cpi\_au_{t}$: The Consumer Price Index, Australia
  +   $unemr\_au_{t}$: The Unemployment rate, Australia
  +   $impor\_au_{t}$: The Imports of Goods and Services, Australia
  +   $expor\_au_{t}$: The Exports of Goods and Services, Australia
* US economic indicators
  +   $crate\_us_{t}$: The Federal Funds Effective Rate, United States
  +   $rgdp\_us_{t}$: The Real Gross Domestic Product, United States
  +   $cpi\_us_{t}$: The Consumer Price Index, United States
  +   $unemr\_us_{t}$: The Unemployment rate, United States
  +   $impor\_us_{t}$: The Imports of Goods and Services, United States
  +   $expor\_us_{t}$: The Exports of Goods and Services, United States
  
### Data Cleaning
For the data cleaning part, most data for AUS is from Reserve Bank of Australia (RBA) and Australian Bureau of Statistics (ABS), and data for the US is from FRED, the dataset spans from 1990 Q1 to 2023 Q4, comprising 136 observations. To better fit the model, the data has been transformed to 'quarter' to ensure that seasonality effects are removed and logged transformations have been applied to most data except exchange rate and cash rate to reducing outlier effects.
```{r, warning=FALSE, message=FALSE}
#| echo: false
#| message: false
#| warning: false
#| results: hide

library(fredr)
library(readrba)
library(readabs)
library(xts)
library(tseries)
library(fUnitRoots) 
fredr_set_key("75b470c4883ecfd5a7b4185f30437bd0")
```

```{r, warning=FALSE, message=FALSE}
#| echo: false
#| message: false
#| warning: false

#Y variable
# AUD/USD exchange rate quarterly
ex_rate <- read_rba(series_id = "FXRUSD")
ex_rate$date <- as.Date(ex_rate$date)
erate <- xts(ex_rate$value, ex_rate$date)
erate <- apply.quarterly(erate,mean)
index(erate) <- seq(as.Date("1969-09-01"), by = "3 months", length.out = nrow(erate))


#gold price (2006-)
#gpricelink <- "https://query1.finance.yahoo.com/v7/finance/download/GC%3DF?period1=1262304000&period2=1703980800&interval=1mo&filter=history&frequency=1mo&includeAdjustedClose=true"
#gprice  <- read.csv(gpricelink)
#gprice$Date <- as.Date(gprice$Date)
#gprice <- xts(gprice$Adj.Close, gprice$Date)
#gprice <- apply.quarterly(gprice,mean)

# Australia real gdp seasonal adjusted quarterly
#rgdp_au <- read_abs(series_id = "A2304404C")  
rgdp_au <- read_rba(series_id = "GGDPCVGDP")
rgdp_au <- xts::xts(rgdp_au$value, rgdp_au$date)
index(rgdp_au)   <- seq(as.Date("1959-09-01"), by = "3 months", length.out = nrow(rgdp_au))

#cash rate/interest rate of AUS quartly
cashrate<- read_cashrate(type = c("target"))
crate_au<- xts(cashrate$value, cashrate$date)
crate_au<- apply.quarterly(crate_au,mean)
crate_au<- xts(crate_au, seq(as.Date("1990-03-01"), by = "quarter", length.out = length(crate_au)))

#CPI quartly
# cpi_au <- read_rba(series_id = "GCPIAG")
cpi_au <- read_abs(series_id = "A2325846C")  
cpi_au <- xts::xts(cpi_au$value, cpi_au$date)

#unemployment rate quartly
#unemprate <-read_rba(series_id = "GLFSURSA")
unemprate <- read_abs(series_id = "A84423050A") 
unemr_au<- xts(unemprate$value, unemprate$date)
unemr_au<- apply.quarterly(unemr_au,mean)

# International Trade in Goods and Services seasonal adjusted_quartly
exportaus <- read_abs(series_id = "A2718603V")   
expor_au<- xts(exportaus $value, exportaus $date)
expor_au<- abs(expor_au)
expor_au<- apply.quarterly(expor_au,mean)

importaus <- read_abs(series_id = "A2718577A")     
impor_au<- xts(importaus$value, importaus$date)
impor_au<- apply.quarterly(impor_au,mean)

# America data
# us gdp
#rgdpus <- fredr(series_id = "A939RX0Q048SBEA")
rgdpus     <- fredr(series_id = "GDPC1")
rgdp_us     <- to.quarterly(xts(rgdpus$value, rgdpus$date), OHLC = FALSE)
index(rgdp_us) <- seq(as.Date("1947-03-01"), by = "3 months", length.out = nrow(rgdp_us))

#Federal Funds Effective Rate/interest rate quartly
usdratelink = "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1318&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=DFF&scale=left&cosd=1954-07-01&coed=2024-03-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Daily%2C%207-Day&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2024-04-01&revision_date=2024-04-01&nd=1954-07-01"
crate_us <- read.csv(usdratelink)
crate_us$DATE <- as.Date(crate_us$DATE)
crate_us <- xts(crate_us$DFF, order.by = crate_us$DATE)
crate_us <- apply.quarterly(crate_us,mean)
crate_us<- xts(crate_us, seq(as.Date("1954-09-01"), by = "quarter", length.out = length(crate_us)))

# cpi quartly
cpiusd  <- fredr(series_id = "USACPIALLMINMEI")
cpi_us<- xts(cpiusd$value, cpiusd$date)
cpi_us<- apply.quarterly(cpi_us,mean)

# unemployment quartly
unemprate_usd = fredr(series_id = "UNRATE")
unemr_us <- xts(unemprate_usd$value, unemprate_usd$date)
unemr_us<- apply.quarterly(unemr_us,mean)

#export_usd——quartly
usdexportink = "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1318&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=EXPGS&scale=left&cosd=1947-01-01&coed=2023-10-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Quarterly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2024-03-30&revision_date=2024-03-30&nd=1947-01-01"
expor_us <- read.csv(usdexportink)
expor_us$DATE <- as.Date(expor_us$DATE)
expor_us <- xts::xts(expor_us$EXPGS, order.by = expor_us$DATE)
index(expor_us) <- seq(as.Date("1947-03-01"), by = "3 months", length.out = nrow(expor_us))

#import_usd_quartly
usdimportlink = "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1318&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=IMPGS&scale=left&cosd=1947-01-01&coed=2023-10-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Quarterly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2024-03-30&revision_date=2024-03-30&nd=1947-01-01"
impor_us <- read.csv(usdimportlink)
impor_us$DATE <- as.Date(impor_us$DATE)
impor_us <- xts::xts(impor_us$IMPGS, order.by = impor_us$DATE)
index(impor_us) <- seq(as.Date("1947-03-01"), by = "3 months", length.out = nrow(impor_us))

```

```{r}
#| echo: false
#| message: false
#| warning: false

# log transformation of data
variables <- c("cpi_au", "cpi_us", "rgdp_au", "rgdp_us", "impor_au", "impor_us", "expor_au", "expor_us")

for(var in variables) {
  assign(var, log(get(var)))
}

#gprice <- log(gprice)
```


```{r}
#| echo: false
#| message: false
#| warning: false
 
# All Variables
merged_data = na.omit(merge(erate, 
                            cpi_au, cpi_us, 
                            crate_au, crate_us, 
                            expor_au, expor_us,  
                            impor_au, impor_us, 
                            rgdp_au, rgdp_us,
                            unemr_au, unemr_us))

# Defining your column name vector:
variable_names <- c("exchange rate", "cpi_au", "cpi_us", 
                    "cashrate_au", "cashrate_us", "export_au", "export_us",
                    "import_au", "import_us", "realgdp_au", "realgdp_us",
                    "unemployemtrate_au", "unemployemtrate_us")


colnames(merged_data)   <- variable_names


```
#### Visualisation of data
Plot the variables to see the patterns of data. It shows from the plots that the exchange rate and cash rate for the US fluctuate over time, the cash rate and unemployment for AU show a downward trend and all other variables have a clear upward trend, with the exports、imports, and GDP for both countries have a clear drop during the COVID-19 period.
```{r all variables plot}
#| echo: false
#| message: false
#| warning: false

par(mfrow = c(4, 4), mar=c(2,2,2,2))
for (i in 1:13) { 
  ts.plot(merged_data[, i], main = colnames(merged_data)[i], 
          ylab = "", xlab = "")
}

```
Since most variables show a non-stationary pattern. To determine whether a unit root is present in a time series dataset, the ADF test will be conducted below.

#### Augmented Dickey-Fuller test for log transformed variables except exchange rates and cash rates.

From the plot we can observe all ACF plots have a high degree of persistence over time, indicating there is significant autocorrelation in the time series data.
```{r}
#| echo: false
#| message: false
#| warning: false

par(mfrow = c(4, 4), mar=c(2,2,2,2))
for (i in 1:13){
acf = acf(merged_data[,i], plot = FALSE)[1:20]
plot(acf, main = "")
title(main = paste(colnames(merged_data)[i]), line = 0.5)
}
```
```{r}
#| echo: false
#| message: false
#| warning: false
adf_test <- list()
for (i in 1:13) {
  adf_result = adf.test(merged_data[,i], k = 4)
  adf_test[[i]] <- adf_result
}
```
Below is the p-value of each variable and only the cash rate for AUS has a p-value less than 0.05 which indicates that $crateau_{t}$ is stationary.
```{r}
#| echo: false
#| message: false
#| warning: false
adf_table <- data.frame(p_value = numeric(length(adf_test)))

for (i in 1:length(adf_test)) {adf_table[i, "p_value"] = round(adf_test[[i]]$p.value,3)
}

rownames(adf_table) <- variable_names

colnames(adf_table)<- c("P-value")
print(adf_table)

```
Below is the ADF test result for all non-stationary data taking the first difference. All variables in the first differences are stationary as the null hypothesis of non-stationary can be rejected.
```{r}
#| echo: false
#| message: false
#| warning: false
#take the first difference
nonstationary_merged_data <- subset(merged_data, select = -c(cashrate_au))

dff_merged_data <- na.omit(nonstationary_merged_data - lag(nonstationary_merged_data))
```

```{r}
#| echo: false
#| message: false
#| warning: false
# ADF test
dff_adf_test <- list()
for (i in 1:12) {
  dff_adf_result = adf.test(dff_merged_data[,i], k = 4)
  dff_adf_test[[i]] <- dff_adf_result
}

# View the ADF test results
dff_adf_table <- data.frame(p_value = numeric(length(dff_adf_test)))

# Fill in the data frame with the test results
for (i in 1:length(dff_adf_test)) {
  dff_adf_table[i, "p_value"] = round(dff_adf_test[[i]]$p.value,3)

}
rownames(dff_adf_table) <- variable_names[-4]


colnames(dff_adf_table)<- c("P-value")
print(dff_adf_table)
```
It can be concluded that all variables are integrated at 1 with $crateau_{t}$ being the only stationary so the first difference should be taken to keep the time series stationary.

# Model and Hypotheses
In this research, the VAR(p) model will be applied to forecast the AUD/USD exchange rate, below is the basic model that is used in this research.

#### The basic VAR(p) model

```{=tex}
\begin{aligned}
 y_{t}=\mu_{0}+A_{1}y_{t-1}+\cdots+A_{p}y_{t-p}+\epsilon_{t} \\
\epsilon_{t}|Y_{t-1}\sim iid(0_{13},\Sigma)
\end{aligned}
```

For time $t$ = 1,2,.....,$T$：

-   $y_t$ is a $N(13)\times 1$ vector of observations at time $t$
-   $\mu_0$ is a $N(13)\times 1$ vector of constant terms
-   $A_i$ is a $N(13)\times N(13)$ matrix of autoregressive slope parameters
-   $\epsilon_t$ is a $N(13)\times 1$ vector of error terms and follows a multivariate white noise process
-   $Y_{t-1}$ is the information set collecting observations on y up to time $t-1$
-   $\Sigma$ is a $N(13)\times N(13)$ covariance matrix of the error term

```{=tex}
\begin{aligned}
y_{t}=\begin{pmatrix}
erate_{t}\\
crate\_au_{t} \\
rgdp\_au_{t}\\
cpi\_au_{t} \\
unemr\_au_{t}\\
impor\_au_{t} \\
expor\_au_{t} \\
crate\_us_{t} \\
rgdp\_us_{t} \\
cpi\_us_{t} \\
unemr\_us_{t} \\
impor\_us_{t}\\
expor\_us_{t}\\

\end{pmatrix}

\end{aligned}
```

#### The Matrix notation
```{=tex}
\begin{align}
Y =XA+E\\
E|X\sim \mathcal{MN}_{T\times N}(0, \Sigma, I_T)\\
Y|X,A,\Sigma\sim \mathcal{MN}_{T\times N} (XA, \Sigma, I_T)
\end{align}
```
Where $Y$ is a $T\times 13$ Matrix, $X$ is a $T\times(1+p\times13)$, $A$ is a $(1+p\times13)\times13$ matrix that contains $\mu_{0}$ and vectors of the autoregressive slope parameters and $E$ is a $T\times13$ matrix contains vetors of error terms.

For further research, we may use the predictive density function like 1-period forecast and forecast with Bayesian VARS.